Calculating the Solvent Exposed Surface Area of Proteins with Intersecting Spheres

### 2020-06-05 Paper.tex
Small Edits

\subsection{Shrake-Rupley Algorithm}

@@

The SASA of an atom with radius $\varrho_i$ lies on a sphere where $K=\frac{1}{\varrho_i^2}$. We can divide this SASA into many polygons bounded by the circles of intersection with neighboring spheres (See Figure). If $M$ is one particular region of that SASA, we can write 
\begin{align}\nonumber
&\frac{1}{\varrho_i^2}\int_M dA + \int_{\partial M} k_g ds= 2\pi \\
&\implies A=\varrho_i^2\left(2\pi-\int_{\partial M} k_g ds\right)
\end{align}

@@

Now we can find the area of both spheres. We know that $A = 4\pi r^2$. For two spheres $A = A_i + A_j$. Recall the area equation stated in the model. For two balls we can say: $A = 4\pi\varrho^2\sum{\sigma_i} + 4\pi\varrho^2\sum{\sigma_j}$. After plugging in $\sigma$ and simplifying the area equation:
\begin{align*}
&4\pi\varrho^2\frac{\varrho_i + \zeta_i}{2\varrho_i} + 4\pi\varrho_j^2\frac{\varrho_j + \zeta_j}{2\varrho_j} \\
&\text{After a few algebraic simplifications:} \\
&2\pi\varrho_i^2 +2\pi\varrho_j^2 +\pi\varrho_i\left(\zeta_{ij} + \frac{\varrho_i^2 - \varrho_j^2}{\zeta_{ij}}\right)+\pi\varrho_j\left(\zeta_{ij} + \frac{\varrho_j^2 - \varrho_i^2}{\zeta_{ij}}\right) \\
&A  = 2\pi*(\varrho_i^2 + \varrho_j^2) + \pi(\varrho_i+\varrho_j)\left[\zeta_{ij}+\frac{(\varrho_i-\varrho_j)^2}{\zeta_{ij}}\right] \\
&\text{Then we find the derivative of the area with respect to $\zeta_{ij}$} \\
&\frac{dA}{d\zeta_{ij}} = \pi(\varrho_i+\varrho_j)\left[1-\frac{(\varrho_i - \varrho_j)^2}{\zeta_{ij}^2}\right] \\
&\text{This is term they get after some simplification.}
\end{align*}

@@

\section{Application of ALPHAMOL to Molecular Modeling}
To test the time cost of \textsc{alphamol} (the computational implement of the method described in $\S3$) we took 250 PDB structures from the RCSB database (listed in the supplemental) and calculated the number of polar atoms buried in the structure that are not making hydrogen bonds. This calculation indirectly uses a SASA calculator to determine if atoms close to the surface are truly buried. If they have any SASA that is nonzero, the atom is assumed to be making a hydrogen bond with the implicit solvent. We compared the time taken to complete the buried unsatisfied polars calculation for the \textsc{alphamol} and Shrake-Rupley SASA calculators. We found that \textsc{alphamol} is slower, but it is more accurate.

@@

@@

\bibitem{Levitt} 
Bryant, R., Edelsbrunner, H., Koehl, P., Levitt, M. The Area Derivative of a Space-Filling Diagram. Discrete \& Computational Geometry. 32 (2004), 293-308.

>>>>

\subsection{The Shrake-Rupley Algorithm}

@@

The SASA of an atom with radius $\varrho_i$ lies on a sphere where $K=\frac{1}{\varrho_i^2}$. We can divide this SASA into many polygons bounded by the circles of intersection with neighboring spheres (See Figure 1). If $M$ is one particular polygon, we can write 
\begin{align}\nonumber
&\frac{1}{\varrho_i^2}\int_M dA + \int_{\partial M} k_g ds= 2\pi \\
&\implies A=\varrho_i^2\left(2\pi-\int_{\partial M} k_g ds\right)
\end{align}

@@

Now we can find the area of both spheres. We know that $A = 4\pi r^2$. For two spheres $A = A_i + A_j$. Recall the area equation stated in the model. For two balls we can say: $A = 4\pi\varrho^2\sum{\sigma_i} + 4\pi\varrho^2\sum{\sigma_j}$. After plugging in $\sigma$ and simplifying the area equation:
\begin{align*}
&4\pi\varrho^2\frac{\varrho_i + \zeta_i}{2\varrho_i} + 4\pi\varrho_j^2\frac{\varrho_j + \zeta_j}{2\varrho_j} \\
&\text{After a few algebraic simplifications:} \\
&2\pi\varrho_i^2 +2\pi\varrho_j^2 +\pi\varrho_i\left(\zeta_{ij} + \frac{\varrho_i^2 - \varrho_j^2}{\zeta_{ij}}\right)+\pi\varrho_j\left(\zeta_{ij} + \frac{\varrho_j^2 - \varrho_i^2}{\zeta_{ij}}\right) \\
&A  = 2\pi(\varrho_i^2 + \varrho_j^2) + \pi(\varrho_i+\varrho_j)\left[\zeta_{ij}+\frac{(\varrho_i-\varrho_j)^2}{\zeta_{ij}}\right] \\
&\text{Then we find the derivative of the area with respect to $\zeta_{ij}$} \\
&\frac{dA}{d\zeta_{ij}} = \pi(\varrho_i+\varrho_j)\left[1-\frac{(\varrho_i - \varrho_j)^2}{\zeta_{ij}^2}\right] \\
&\text{This is term they get after some simplification.}
\end{align*}

@@

\section{Application of ALPHAMOL to SASA Calculations}
To test the time cost of \textsc{alphamol} (the computational implement of the method described in $\S3$) we took 250 PDB structures from the RCSB database (listed in the supplemental) and calculated the number of polar atoms buried in the structure that are not making hydrogen bonds. This calculation indirectly uses a SASA calculator to determine if atoms close to the surface are truly buried. If they have any SASA that is nonzero, the atom is assumed to be making a hydrogen bond with the implicit solvent. We compared the time taken to complete the buried unsatisfied polars calculation for the \textsc{alphamol} and Shrake-Rupley SASA calculators. We found that \textsc{alphamol} is slower, but it is more accurate.

@@

\bibitem{Bryant} 
Bryant, R., Edelsbrunner, H., Koehl, P., Levitt, M. The Area Derivative of a Space-Filling Diagram. Discrete \& Computational Geometry. 32 (2004), 293-308.

@@
